# Clustering and Classification

## Analysis task 2
```{r}
library(MASS)
Boston
head(Boston,2)
str(Boston)
``` 
The data has 506 rows (observations) and 14 columns (variables). All variables are numeric The file contains information about housing values in suburbs of Boston. More information is available here:[link] (https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/Boston.html)

## Analysis task 3
```{r}  
summary(Boston)
Boston_cor <-cor(Boston)
library(corrplot)
corrplot(Boston_cor, method="circle")

library(purrr)
library(tidyr)
library(ggplot2)

Boston %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()
par(mar=c(1,1,1,1))
pairs(Boston)


```
The strongest positive correlations can be found between tax and rad and the strongest negative between dis and nox, dis and indus, dis and age as well as medv and lstat. Looking at the diskription of the variables (see link below), the relationships make sense.

Concerning the distributions of values: age and black are skewed to higher values, chas, tax and rad have two groups of values, rm and medv have sort of a normal distribution and dis and crim are skewed to low values.


## Analysis task 4
```{r}  
boston_scaled <- scale(Boston)
summary(boston_scaled)
boston_scaled_df <- as.data.frame( boston_scaled)
sd(boston_scaled_df$zn)

bins <- quantile(boston_scaled)
crime <- cut(boston_scaled_df$crim, breaks = bins, include.lowest = TRUE, labels = c("low", "med_low", "med_high", "high"))
head(crime,2)

boston_scaled_df$crimeClass<-crime
head(boston_scaled_df,2)
colnames(boston_scaled_df)
boston_scaled_df <- boston_scaled_df[,c("zn","indus","chas","nox","rm","age","dis","rad","tax","ptratio","black","lstat","medv","crimeClass")]
head(boston_scaled_df,2)
colnames(boston_scaled_df)

n <- nrow(boston_scaled_df)

ind <- sample(n,  size = n * 0.8)

train_boston <- boston_scaled[ind,]
test_boston <- boston_scaled[-ind,]

```


## Analysis task 5
```{r}  


```



## Analysis task 6
```{r} 

```

## Analysis task 7
```{r} 

```

